Reproducible, flexible and high throughput data extraction from primary literature: The \pkg{metaDigitise} \proglang{R} package \vskip15pt Supplementary Materials

### 2018-02-28 docs/MEE_submission/manuscript_MEE.tex
Added small changes to MS, 3100 words

%% - \Abstract{} almost as usual
\section*{Abstract}
1. Research synthesis requires data extraction from primary studies with effect sizes for meta-analyses being calculated from summary statistics. However, exact values of such statistics are commonly hidden in figures. 
2. Extracting summary statistics from figures can be a slow process that is not easily reproducible. Additionally, current software lacks an ability to incorporate important meta-data (e.g., sample sizes, treatment names) about experiments. Furthermore, extraction software is often independent of analytical software preventing the automatic calculation of effect size statistics and making it more challenging to implement user defined analytical pipelines.
3. Here we present the R package \pkg{metaDigitise} which extracts descriptive statistics such as means, standard deviations and correlations from the four plot types: 1) mean/error plots (e.g. bar graphs with standard errors), 2) box plots, 3) scatter plots and 4) histograms. \pkg{metaDigitise} is user-friendly and easy to learn as it interactively guides the user through the data extraction process. Notably, it enables large-scale extraction by automatically loading image files, letting the user stop processing, edit and add to the resulting data fame at any point. 
4. Digitised data can be easily re-plotted and checked, facilitating reproducible data extraction from plots with little inter-observer bias. Such approaches allow collaborative extraction of data. We hope that by making the process of figure extraction more flexible and easy to conduct it will improve transparency and quality of meta-analyses in the future.
%% 230 words
 Keywords: meta-analysis, comparative analysis, data extraction, \proglang{R}, reproducibility, figures, images, summary statistics

@@

%% 414 words
%%%---------------------------------
%%%---------------------------------
%%%---------------------------------

@@

%% 779 words

@@

%% 183 words

@@

%% 577 words

@@

%% 275 words

@@

\section*{Acknowledgments}
We thank the I-DEEL group and colleagues at UNSW for for testing, providing feedback and digitising including: Rose O'Dea, Fonti Kar, Malgorzata Lagisz, Julia Riley, Diego Barneche, Erin Macartney, Ivan Beltran, Gihan Samarasinghe, Dax Kellie, Jonathan Noble, Yian Noble and Alison Pick. J.L.P. was supported by a Swiss National Science Foundation Early Mobility grant (P2ZHP3\_164962), D.W.A.N. was supported by an Australian Research Council Discovery Early Career Research Award (DE150101774) and UNSW Vice Chancellors Fellowship and S.N. an Australian Research Council Future Fellowship (FT130100268).

>>>>

%% - \Abstract{} almost as usual
\section*{Abstract}
1. Research synthesis requires data extraction from primary studies with effect sizes for meta-analyses being calculated from summary statistics. However, exact values of such statistics are commonly hidden in figures. 
2. Extracting summary statistics from figures can be a slow process that is not easily reproducible. Additionally, current software lacks an ability to incorporate important meta-data (e.g., sample sizes, treatment / variable names) about experiments and is not integrated with other software to streamline analysis pipelines.
3. Here we present the R package \pkg{metaDigitise} which extracts descriptive statistics such as means, standard deviations and correlations from the four plot types: 1) mean/error plots (e.g. bar graphs with standard errors), 2) box plots, 3) scatter plots and 4) histograms. \pkg{metaDigitise} is user-friendly and easy to learn as it interactively guides the user through the data extraction process. Notably, it enables large-scale extraction by automatically loading image files, letting the user stop processing, edit and add to the resulting data fame at any point. 
4. Digitised data can be easily re-plotted and checked, facilitating reproducible data extraction from plots with little inter-observer bias. We hope that by making the process of figure extraction more flexible and easy to conduct it will improve the transparency and quality of meta-analyses in the future.
%% 209 words
 Keywords: meta-analysis, comparative analysis, data extraction, \proglang{R}, reproducibility, figures, images, summary statistics

@@

%% 410 words
%%%---------------------------------
%%%---------------------------------
%%%---------------------------------

@@

%% 777 words

@@

%% 171 words

@@

%% 571 words

@@

%% 271 words

@@

% 68
\section*{Acknowledgments}
We thank the I-DEEL group and colleagues at UNSW for for testing, providing feedback and digitising including: Rose O'Dea, Fonti Kar, Malgorzata Lagisz, Julia Riley, Diego Barneche, Erin Macartney, Ivan Beltran, Gihan Samarasinghe, Dax Kellie, Jonathan Noble, Yian Noble and Alison Pick. J.L.P. was supported by a Swiss National Science Foundation Early Mobility grant (P2ZHP3\_164962), D.W.A.N. was supported by an Australian Research Council Discovery Early Career Research Award (DE150101774) and UNSW Vice Chancellors Fellowship and S.N. an Australian Research Council Future Fellowship (FT130100268).
